---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sotoncolours)
```

# University of Southampton Colour Palettes

`sotoncolours` provides colour palettes for use with `plot` or `ggplot2`
based upon the [University of Southampton colour palettes](https://www.southampton.ac.uk/brand/category/colour/) of  `neutral`,
`marine`, `horizon` and `web`.

The code is largely re-purposed from the [ochRe package](https://github.com/ropenscilabs/ochRe)

It has two functions for use with `ggplot2`, either as discrete colours or gradients:

+ `scale_fill_uos()`
+ `scale_colour_uos()`

and the palettes or specific colours can be called using `uos_cols()` and/or 
`uos_colours`. Colours can be used in the base plotting functions too. 

See examples below and more details in the vignette.

## Installation

Use can install from github using `devtools`. Run `install.packages("devtools")`
if necessary first.

```{r installation, eval=FALSE}
devtools::install_github("ab604/sotoncolours", build_vignettes = TRUE)
```

```{r vignette,eval=FALSE}
vignette("sotoncolours")
```


and then load the library as usual:

```{r load-lib, eval=FALSE}
library(sotoncolours)
```

## Colour and palette choices

The package contains 19 colours and 7 colour palettes contained within 
a named character vector called `uos_colours`:

```{r colours}
# Named character vector of the colours
uos_colours
```


```{r palettes, echo=TRUE}
# See names of all palettes
names(uos_palettes)
```

## The palettes

```{r show_palettes}
pal_names <- names(uos_palettes)

par(mfrow=c(length(uos_palettes)/2, 2), lheight = 2, mar=rep(1, 4), adj = 0)
for (i in 1:length(uos_palettes)){
    viz_palette(uos_palettes[[i]], pal_names[i])
}
```

## Example

Here is an example using [Alison Horst's Palmer Penguins dataset](https://github.com/allisonhorst/palmerpenguins)
and `ggplot2` and `scale_fill_uos()`

```{r penguin-example, warning=FALSE,message=FALSE}
library(palmerpenguins)
library(ggplot2)

ggplot(data = penguins, aes(x = flipper_length_mm)) +
  geom_histogram(aes(fill = species), 
                 alpha = 0.5, 
                 position = "identity") +
  scale_fill_uos(palette = "contrast1") +
  theme_minimal() +
  labs(x = "Flipper length (mm)",
       y = "Frequency",
       title = "Penguin flipper lengths")
```

Here is a base R `plot` using the `horizon` palette directly via the `colorRampPalette()` function.

```{r base-plot, height = 8}
pal <- colorRampPalette(uos_palettes[["horizon"]])
image(volcano, col = pal(20))
```

Individual palettes can be visualised using the `viz_palette` function

```{r}
viz_palette(uos_palettes[["marine"]])
```

It's possible, but messy, to pick any colour combination from `uos_colours`,
but the syntax requires extracting the hex codes using `unname()` and subsetting
`uos_cols` using a character vector of the desired colour names.

It may be easier to create an object for any bespoke palette and pass then
pass it to the respective `ggplot` function.

Here's an example with the Palmer penguin data using
`unname(uos_colours[c("black","shamrock","horizon4")])` as described 
above by creating  a new palette called `my_cols` that is passed to the
`values` argument of `scale_colour_manual()`.

```{r maps, echo=TRUE, message=FALSE}
# Create a new vector of colours to pass to scale_colour manual
# Use unname() to get hex codes from uos_cols()
my_cols <- unname(uos_colours[c("black","shamrock","horizon4")])

ggplot(data = penguins, aes(x = species, y = flipper_length_mm)) +
  geom_boxplot(aes(color = species), width = 0.3, show.legend = FALSE) +
  geom_jitter(aes(color = species), alpha = 0.7, show.legend = FALSE, 
              position = position_jitter(width = 0.2, seed = 0)) +
  scale_colour_manual(values = my_cols) + # pass my_cols to scale_colour_manual
  theme_minimal() +
  labs(x = "Species",
       y = "Flipper length (mm)")
```
