---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sotoncolours)
```

# University of Southampton Colour Palettes

`sotoncolours` provides colour palettes for use with `plot` or `ggplot2`
based upon the [University of Southampton colour palettes](https://www.southampton.ac.uk/brand/category/colour/) of  `neutral`,
`marine`, `horizon` and `web`.

The code is largely re-purposed from the [ochRe package](https://github.com/ropenscilabs/ochRe)

## The palettes

```{r show_palettes, fig.height= 8}
pal_names <- names(uos_palettes)

par(mfrow=c(length(uos_palettes)/2, 2), lheight = 2, mar=rep(1, 4), adj = 0)
for (i in 1:length(uos_palettes)){
    viz_palette(uos_palettes[[i]], pal_names[i])
}
```

## Example

Here is an example using [Alison Horst's Palmer Penguins dataset](https://github.com/allisonhorst/palmerpenguins)
and `ggplot2`

```{r penguin-example, warning=FALSE,message=FALSE}
library(palmerpenguins)
library(ggplot2)

ggplot(data = penguins, aes(x = flipper_length_mm)) +
  geom_histogram(aes(fill = species), 
                 alpha = 0.5, 
                 position = "identity") +
  scale_fill_uos(palette = "contrast1") +
  theme_minimal() +
  labs(x = "Flipper length (mm)",
       y = "Frequency",
       title = "Penguin flipper lengths")
```

Here is a base R `plot` using the `horizon` palette directly via the `colorRampPalette()` function.

```{r base-plot}
pal <- colorRampPalette(uos_palettes[["horizon"]])
image(volcano, col = pal(20))
```

Individual palettes can be visualised using the `viz_palette` function

```{r}
viz_palette(uos_palettes[["marine"]])
```
