---
title: "Using the University of Southampton colour palettes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using the University of Southampton colour palettes}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(fig.width = 7, 
                      echo = FALSE, 
                      warning = FALSE, 
                      message=FALSE)
```

```{r}
library(tidyverse)
library(sotoncolours)
library(dichromat)
library(colorspace)
library(palmerpenguins)
library(raster)
library(gridExtra)
library(patchwork)
```

# Introduction

The [**sotoncolours**](https://github.com/ab604/sotoncolours) package contains colour palettes based on the University of Southampton colour scheme.

# Installation

```{r install, eval=FALSE}
devtools::install_github("ab604/sotoncolours")
```

# Colour and palette choices

The package contains 19 colours and 7 colour palettes.

```{r colours}
# Names of the colours
names(uos_cols())
```


```{r palettes, echo=TRUE}
# See names of all palettes
names(uos_palettes)
```

To see the colours in a palette by name

```{r palette-colour, echo=TRUE}
# See names of all palettes
names(uos_palettes$contrast1)
```

```{r see-palettes, fig.height = 8}
pal_names <- names(uos_palettes)
par(mfrow=c(length(uos_palettes)/2, 2), lheight = 2, mar=rep(1, 4), adj = 0)
for (i in 1:length(uos_palettes)){
    viz_palette(uos_palettes[[i]], pal_names[i])
}
```


## Colour vision impairment
These plots show how each of the palettes perform for different types of colour vision impairment (using dichromat).

### Green-Blind (Deuteranopia)
```{r dichromate_deat, fig.height = 5}
par(mfrow=c(length(uos_palettes)/2, 2), lheight = 2, mar=rep(1, 4), adj = 0)
for (i in 1:length(uos_palettes)){
    viz_palette(dichromat(uos_palettes[[i]], "deutan"), pal_names[i])
}
```


### Red-Blind (Protanopia)
```{r dichromate_protan, fig.height = 5}
par(mfrow=c(length(uos_palettes)/2, 2), lheight = 2, mar=rep(1, 4), adj = 0)
for (i in 1:length(uos_palettes)){
    viz_palette(dichromat(uos_palettes[[i]], "protan"), pal_names[i])
}
```


### Blue-Blind (Tritanopia)
```{r dichromate_tritan, fig.height = 5}
par(mfrow=c(length(uos_palettes)/2, 2), lheight = 2, mar=rep(1, 4), adj = 0)
for (i in 1:length(uos_palettes)){
    viz_palette(dichromat(uos_palettes[[i]], "tritan"), pal_names[i])
}
```


### Desaturated
```{r desaturate, fig.height = 5}
par(mfrow=c(length(uos_palettes)/2, 2), lheight = 2, mar=rep(1, 4), adj = 0)
for (i in 1:length(uos_palettes)){
    viz_palette(desaturate(uos_palettes[[i]]), pal_names[i])
}
```

# Usage

Colours can be called directly using `uos_cols()` with a character argument
for the colour name e.g. `"horizon5"`.

```{r}
ggplot(mtcars, aes(hp, mpg)) +
    geom_point(color = uos_cols("horizon5"),
               size = 4, alpha = .8)
```

The package contains color scale functions for **ggplot2**
plots: `scale_color_uos()` and `scale_fill_uos()` calling one of the palettes. 

Here the `marine` and `horizon5` palettes are used for a scatter plot of `mtcars` data.

```{r mtcars, echo = TRUE, message = FALSE}
# Use the marine palette for colours
p1 <- ggplot(mtcars, aes(mpg, wt)) + 
   geom_point(aes(colour = factor(cyl)), size=2, alpha=0.7) + 
   scale_colour_uos(palette="marine") +
    theme_bw() + 
    theme(aspect.ratio=1, legend.position="bottom")

# Use the horizon palette for colours
p2 <- ggplot(data = mpg) + 
   geom_point(mapping = aes(x = displ, y = hwy, color = class), alpha=0.7) +
   scale_colour_uos(palette="horizon") +
    theme_bw() + 
    theme(aspect.ratio=1, legend.position="bottom")

p1 + p2
```

Here the default palette `marine` is used in a stacked bar chart using the `diamonds` data. 

```{r diamonds, echo=TRUE, message=FALSE}
 ggplot(diamonds) + 
  geom_bar(aes(x = cut, fill = clarity)) +
  scale_fill_uos() +  # defaults to marine
  theme_bw() 
```

It's possible, but messy, to pick any colour combination using `uos_cols()`,
but the syntax requires extracting the hex colour values using `unname()` and
passing a vector of colour names as the index to `uos_cols()`

Here's an example with the Palmer penguin data using
`uos_cols()` as described above by creating a new palette called `my_cols` 
that is passed to `scale_colour_manual()`.

```{r maps, echo=TRUE, message=FALSE}
# Create a new vector of colours to pass to scale_colour manual
# Use unname() to get hex codes from uos_cols()
my_cols <- unname(uos_cols()[c('black','shamrock','horizon4')])

ggplot(data = penguins, aes(x = species, y = flipper_length_mm)) +
  geom_boxplot(aes(color = species), width = 0.3, show.legend = FALSE) +
  geom_jitter(aes(color = species), alpha = 0.7, show.legend = FALSE, 
              position = position_jitter(width = 0.2, seed = 0)) +
  scale_colour_manual(values = my_cols) + # pass my_cols to scale_colour_manual
  theme_minimal() +
  labs(x = "Species",
       y = "Flipper length (mm)")
```

Here is a base R `plot` using the `horizon` palette directly via the `colorRampPalette()` function.

```{r base-plot, echo=TRUE}
pal <- colorRampPalette(uos_palettes[["horizon"]])
image(volcano, col = pal(20))
```

Here is the same plot, but by creating a new palette from specific colours
as with the penguin data.

```{r base-plot-2, echo=TRUE}
# Create new palette
my_cols <- unname(uos_cols()[c('marine1','shamrock','horizon3')])

pal <- colorRampPalette(my_cols)
image(volcano, col = pal(20))
```
